---
title: "Lab5"
author: "Emily Tipton"
date: '`r Sys.Date()`'
output: 
  html_document:
    toc: yes
    toc_float: yes
---

# Tasks

## Task 1

```{r}
getwd()
```

## Task 2

### mybin()

```{r}
mybin=function(iter=100,n=10, p=0.5)
  { 
# make a matrix to hold the samples
#initially filled with NA's
sam.mat=matrix(NA,nr=n,nc=iter, byrow=TRUE)
#Make a vector to hold the number of successes in each trial
succ=c()

for( i in 1:iter){
#Fill each column with a new sample
sam.mat[,i]=sample(c(1,0),n,replace=TRUE, prob=c(p,1-p))

#Calculate a statistic from the sample (this case it is the sum)
succ[i]=sum(sam.mat[,i])
}

#Make a table of successes
succ.tab=table(factor(succ,levels=0:n))

#Make a barplot of the proportions
barplot(succ.tab/(iter), col=rainbow(n+1), main="Binomial simulation", xlab="Number of successes")
succ.tab/iter
}
```

### iter = 100
```{r}
iter100 = mybin(iter=100, n=10, p=0.7)
```

### iter = 200
```{r}
iter200 = mybin(iter=200, n=10, p=0.7)
```

### iter = 500
```{r}
iter500 = mybin(iter=500, n=10, p=0.7)
```

### iter = 1000
```{r}
iter1000 = mybin(iter=1000, n=10, p=0.7)
```

### iter = 10000
```{r}
iter10000 = mybin(iter=10000, n=10, p=0.7)
```

### iter10000 table

```{r}
iter10000
```

### Check with dbinom()

```{r}
dbin = dbinom(0:10, 10, 0.7)
names(dbin) = 0:10
dbin
```

The values from the table are approximately the same as the values from the dbinom() function. 

## Task 3

### Sample n=5 without replacement

```{r}
sample(c(1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 0, 0, 0, 0, 0, 0, 0, 0), size = 5, replace = FALSE, prob = NULL)
```

### Sample with replace

```{r}
sample(c(1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 0, 0, 0, 0, 0, 0, 0, 0), size = 5, replace = TRUE, prob = NULL)
```

### myhyper

```{r}
myhyper=function(iter=100,N=20,r=12,n=5){

  # make a matrix to hold the samples
#initially filled with NA's
sam.mat=matrix(NA,nr=n,nc=iter, byrow=TRUE)

#Make a vector to hold the number of successes over the trials
succ=c()
for( i in 1:iter){

  #Fill each column with a new sample
sam.mat[,i]=sample(rep(c(1,0),c(r,N-r)),n,replace=FALSE)

#Calculate a statistic from the sample (this case it is the sum)
succ[i]=sum(sam.mat[,i])
}

#Make a table of successes
succ.tab=table(factor(succ,levels=0:n))

#Make a barplot of the proportions
barplot(succ.tab/(iter), col=rainbow(n+1), main="HYPERGEOMETRIC simulation", xlab="Number of successes")
succ.tab/iter
}
```

### iter = 100

```{r}
hyper100 = myhyper(iter = 100, N=20, r=12, n=5)
```

### iter = 200

```{r}
hyper200 = myhyper(iter = 200, N=20, r=12, n=5)
```

### iter = 500

```{r}
hyper500 = myhyper(iter = 500, N=20, r=12, n=5)
```

### iter = 1000

```{r}
hyper1000 = myhyper(iter = 1000, N=20, r=12, n=5)
```

### iter = 10000

```{r}
hyper10000 = myhyper(iter = 10000, N=20, r=12, n=5)
```

### hyper10000 table

```{r}
hyper10000
```

### Check with dhyper()

```{r}
dhyp = dhyper(0:5, m=12, n=8, k=5)
names(dhyp) = 0:5
dhyp
```

The values from the table are approximately the same as the values from the dhyper() function. 

## Task 4
